Supplemental Information: Aligning marine species range data to better serve science and conservation

### 2016-09-12 ms_4_refs.Rmd
updated formats incl. references, figures, etc; split SI and figures into separate docs.  Also added bit about OBIS/GBIF
finished up draft; working on suppl figs

# Acknowledgments
<!-- Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed. -->

@@

Kaschner K et al. (2015) AquaMaps: Predicted range maps for aquatic species. World wide web electronic publication, www.aquamaps.org, Version 08/2015. <!--___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___-->

IUCN (2015). The IUCN Red List of Threatened Species. Version 2015-4. <http://www.iucnredlist.org>. Downloaded on 21 December 2015.  <!--___since this is not a peer-reviewed publication, can we include this for PNAS reference list?___-->

Halpern BS et al. (2012). An index to assess the health and benefits of the global ocean. Nature, 488(7413), 615-620. <!--_doi:10.1038/nature11397._-->

Halpern BS et al. (2015). Patterns and emerging trends in global ocean health. PloS one, 10(3), e0117863. <!--___doi?___-->

Klein CJ et al. (2015). Shortfalls in the global protected area network at representing marine biodiversity. Scientific Reports 5: 17539. <!--_doi:10.1038/srep17539._-->

Selig ER et al. (2013). Assessing global marine biodiversity status within a coupled socio-ecological perspective. PloS one, 8(4), e60284. <!--_doi:10.1371/journal.pone.0060284._-->

Coll M et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. <!--_used AquaMaps to predict Med biodiversity.  Also: Threshold = 0._-->

*Martin CS et al. (2014). Manual of marine and coastal datasets of biodiversity importance. May 2014 release. Cambridge (UK): UNEP World Conservation Monitoring Centre. 28 pp. <!--(+ 4 annexes totalling 174 pp. and one e-supplement). _review of marine data sets and data gaps etc, incl both IUCN and AM as well as many others_-->

Pimm SL et al. (2014). The biodiversity of species and their rates of extinction, distribution, and protection. Science, 344(6187), 1246752. <!--_uses range maps to show biodiversity areas; may use IUCN range maps.  Also discusses gaps and possible things that can be done about them._-->

Kaschner K, Tittensor DP, Ready J, Gerrodette T, Worm B (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. <!--_just what the title says - AM development, presence threshold 60%, also analyzes richness as a function of threshold_-->

Molinos JG et al. (2015). Climate velocity and the future global redistribution of marine biodiversity. Nature Climate Change. <!--_doi:10.1038/nclimate2769._-->

Rondinini C, Wilson KA, Boitani L, Grantham H, & Possingham HP (2006). Tradeoffs of different types of species occurrence data for use in systematic conservation planning. Ecology letters, 9(10), 1136-1145. <!--_compares point locality, range maps, and distribution models in terms of omission and commission errors; also outlines Extent of Occurrence and Area of Occupancy distinctions._-->

Jetz W, Sekercioglu CH, and Watson JE (2008). Ecological correlates and conservation implications of overestimating species geographic ranges. Conservation Biology, 22(1), 110-119. <!--_EOO maps are usually highly interpolated and overestimate small-scale occurrence, which may bias research outcomes_-->

Kaschner K, Watson R, Trites AW, Pauly D (2006). Mapping world-wide distributions of marine mammal species using a relative environmental suitability (RES) model.  Marine Ecology Progress Series 316: 285–310. <!--_check this citation journal name..._ _This outlines the basic RES methodology - AM development_-->

Ready J et al. (2010). Predicting the distributions of marine organisms at the global scale. Ecological Modelling 221(3): 467-478. <!--_Presents AM; assessing AquaMaps against other presence-only species models_-->

Spalding MD et al. (2007). Marine ecoregions of the world: a bioregionalization of coastal and shelf areas. BioScience, 57(7), 573-583.

QGIS Development Team (2015). QGIS Geographic Information System. Open Source Geospatial Foundation Project. http://qgis.osgeo.org

IUCN and UNEP-WCMC (2014). The World Database on Protected Areas (WDPA) [On-line]. Cambridge, UK: UNEP- WCMC. Available at: www.protectedplanet.net [Accessed December 1, 2014].

R Core Team (2015). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

>>>>

# Acknowledgments

<!-- Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed. -->

We are thankful to Melanie Frazier, Julia Stewart Lowndes, Halley Froelich, Claire Runge, and two anonymous reviewers for their insightful comments on earlier versions of this manuscript, and to the National Center for Ecological Analysis and Synthesis for computation support.  We acknowledge financial support from the Gordon and Betty Moore Foundation.

-----

@@

### 2016-09-12 ms_6_supplemental.Rmd
finished up draft; working on suppl figs

methods:
IUCN raw -> grid -> raster -> final
AM raw -> presence

results/discussion:
quad plot with risk and four species (?)
four representative maps

coral quadrant shifts

AM thresholds? not mentioned in body of paper as of yet

@@

__Fig S1.__  Examples of AquaMaps and IUCN raw species range data. (A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) (AquaMaps 2013e) showing global range with varying probabilities of occurrence assigned to 0.5° grid cells (A) and presence within 0.5° grid cells (B). (C, D) IUCN distribution of _T. alalunga_ (IUCN 2011b) represented as extent of occurrence polygons (C) and presence within 0.5° grid cells (D).

<!-- 	
Fis-22832, Thunnus alalunga, 21856, citation general IUCN; 
Reviewed distribution maps for Thunnus alalunga (Albacore), with modelled year 2100 native range map based on IPCC A2 emissions scenario. www.aquamaps.org, version of Aug. 2013. Web. Accessed 11 Sep. 2016.
-->

__Fig S2.__ Rasterizing shapefiles provided by IUCN. A portion of the _T. alalunga_ range map is used to exemplify the rasterization process.  To enable direct comparison of IUCN species ranges to AquaMaps species ranges, the raw IUCN polygon (A) is overlaid with a 0.5° degree grid matching the AquaMaps grid (B).  Each cell is assigned a value of "present" if the cell overlaps any portion of the polygon (C).  The resulting raster (D).

@@

__Fig S3.__ Representative species maps (AquaMaps 2013abcd; Conch 2013ab; IUCN 2011a; IUCN 2012) to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.
(A) Distribution-matched: _Conus episcopatus_, the dignified cone snail.  Distributions show excellent overlap in the western Pacific, though IUCN range extends well beyond the bounds of the AquaMaps range.
(B) Well-matched: _Kajikia albida_, the Atlantic white marlin.  Distributions from each data set show nearly complete overlap, and very similar range size.
(C) Poorly-matched: _Acanthurus nigroris_, the blue-lined surgeonfish.  IUCN predicts species distribution only near the Hawaiian islands; AquaMaps predicts extensive distribution throughout the central and western Pacific Ocean.  The datasets align in neither distribution nor range size.
(D) Area-matched: _Conus magnificus_, the magnificent cone snail.  Distributions overlap in the southern Pacific, but align poorly elsewhere.  The range sizes are similar.

<!--
W-Msc-215455, Conus episcopatus, 192664 citation Conch Books, Hackenheim, Germany; Computer generated distribution maps for Conus episcopatus, with modelled year 2100 native range map based on IPCC A2 emissions scenario. www.aquamaps.org, version of Aug. 2013. Web. Accessed 11 Sep. 2016.

W-Msc-215551, Conus magnificus, 192573 citation Conch Books, Hackenheim, Germany; Computer generated distribution maps for Conus magnificus, with modelled year 2100 native range map based on IPCC A2 emissions scenario. www.aquamaps.org, version of Aug. 2013. Web. Accessed 11 Sep. 2016.

Fis-31079, Acanthurus nigroris, 177999 citation IUCN; Reviewed distribution maps for Acanthurus nigroris (Bluelined surgeonfish), with modelled year 2100 native range map based on IPCC A2 emissions scenario. www.aquamaps.org, version of Aug. 2013. Web. Accessed 11 Sep. 2016.

Fis-164253, Kajikia albida, 170322 citation IUCN; Computer generated distribution maps for Kajikia albida (Atlantic white marlin), with modelled year 2100 native range map based on IPCC A2 emissions scenario. www.aquamaps.org, version of Aug. 2013. Web. Accessed 11 Sep. 2016.
-->

>>>>

<!-- methods: -->
<!-- IUCN raw -> grid -> raster -> final -->
<!-- AM raw -> presence -->

<!-- results/discussion: -->
<!-- quad plot with risk and four species (?) -->
<!-- four representative maps -->

<!-- coral quadrant shifts -->

<!-- AM thresholds? not mentioned in body of paper as of yet -->

@@

![](figures/Fig_S1ab.png)

![](figures/Fig_S1cd.png)

> __Fig S1.__  Examples of AquaMaps and IUCN raw species range data. (A, B) AquaMaps distribution of _Thunnus alalunga_ (Albacore Tuna) (AquaMaps 2013e) showing global range with varying probabilities of occurrence assigned to 0.5° grid cells (A) and presence within 0.5° grid cells (B). (C, D) IUCN distribution of _T. alalunga_ (IUCN 2011b) represented as extent of occurrence polygons (C) and presence within 0.5° grid cells (D).

<!-- Fis-22832, Thunnus alalunga, 21856 -->

![](figures/Fig_S2a.png)

![](figures/Fig_S2b.png)

![](figures/Fig_S2c.png)

![](figures/Fig_S2d.png)

> __Fig S2.__ Rasterizing shapefiles provided by IUCN. A portion of the _T. alalunga_ range map is used to exemplify the rasterization process.  To enable direct comparison of IUCN species ranges to AquaMaps species ranges, the raw IUCN polygon (A) is overlaid with a 0.5° degree grid matching the AquaMaps grid (B).  Each cell is assigned a value of "present" if the cell overlaps any portion of the polygon (C).  The resulting raster (D).

@@

![](figures/Fig_S3.png)

> __Fig S3.__ Representative species maps (AquaMaps 2013abcd; Conch 2013ab; IUCN 2011a; IUCN 2012) to illustrate each quadrant from Fig 2A.  Each map is positioned to match its quadrant in Fig 2A.  FAO Major Fishing Area boundaries [@fao_fisheries_and_aquaculture_department_cwp_2015] are outlined in light grey.
(A) Distribution-matched: _Conus episcopatus_, the dignified cone snail.  Distributions show excellent overlap in the western Pacific, though IUCN range extends well beyond the bounds of the AquaMaps range.
(B) Well-matched: _Kajikia albida_, the Atlantic white marlin.  Distributions from each data set show nearly complete overlap, and very similar range size.
(C) Poorly-matched: _Acanthurus nigroris_, the blue-lined surgeonfish.  IUCN predicts species distribution only near the Hawaiian islands; AquaMaps predicts extensive distribution throughout the central and western Pacific Ocean.  The datasets align in neither distribution nor range size.
(D) Area-matched: _Conus magnificus_, the magnificent cone snail.  Distributions overlap in the southern Pacific, but align poorly elsewhere.  The range sizes are similar.

<!--
W-Msc-215455, Conus episcopatus, 192664
W-Msc-215551, Conus magnificus, 192573
Fis-31079, Acanthurus nigroris, 177999
Fis-164253, Kajikia albida, 170322
-->

## Shifts in coral species range map alignment due to depth clipping

![](figures/Fig_S4.png)

> __Fig. S4.__  Shift in alignment of paired-map coral species due to clipping IUCN ranges to areas shallower than 200 m.  The grey lines represent the change in apparent alignment for a single species.  Most coral species shift rightward from the upper left quadrant to the upper right, with little if any loss in distribution alignment, since in general, only unsuitable habitat has been removed.  Leftward shifts can be seen in species whose larger original range is represented in AquaMaps; by trimming IUCN ranges, the area ratio becomes smaller.
